# Financial Contributions to Presidential Campaigns  
####(Ohio State)
**Dataset:** Financial Contributions to Presidential Campaigns (Ohio State)  
**Time:** 2016  
**The reason to choose this dataset:**    
Ohio is known as a swing state which could forecast the election result by the status of Ohio state.


#### General R library and data loading & enrichment

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r Loading Library}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Libraries for ploting data on the map
#install.packages("ggmap")
#install.packages("ggmap", type = "source")
#install.packages("devtools")
library(devtools)
#install_github("dkahle/ggmap")
#install.packages("zipcode")
library(plyr)

library(ggmap)
library(zipcode)


# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.
library(ggplot2)
library(dplyr)
library(gridExtra)
library(gtable)
library(grid)



# install these if necessary
#install.packages('GGally')
#install.packages('scales')
#install.packages('lattice')
#install.packages('MASS')

# load the ggplot graphics package and the others
library(GGally)
library(scales)
library(memisc)


# Load the package for loading online database
library(data.table)
library(stringr)

library(reshape2)

# Load the package for gender data
library(gender)

# Correlation pakage to replace ggpairs
#install.packages("corrplot")
library(corrplot)

```

```{r  Function: Party categorize function and Get_legend function}
rnParty <- function(x) {
    if(is.na(x)){
      NA
    }else if(x == "Trump, Donald J." | 
             x == "Santorum, Richard J." | 
             x == "Cruz, Rafael Edward 'Ted'" |
             x == "Walker, Scott" |
             x == "Bush, Jeb" |
             x == "Rubio, Marco" |
             x == "Kasich, John R." |
             x == "Carson, Benjamin S."|
             x == "Paul, Rand"|
             x == "Santorum, Richard J."|
             x == "Fiorina, Carly"|
             x == "Jindal, Bobby"|
             x == "Huckabee, Mike"|
             x == "Graham, Lindsey O."|
             x == "Christie, Christopher J."|
             x == "Perry, James R. (Rick)"|
             x == "Pataki, George E."){
      "Republican"
    }else if(x == "Clinton, Hillary Rodham" | x == "Kaine, Tim"| 
             x == "Lessig, Lawrence"| x == "O'Malley, Martin Joseph"){
      "Democratic"
    }else{
      "Other"
    }
}

get_legend <- function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}


```

```{r Load_the_Data}
# Load the Data
oh_data <- read.csv(file = 'P00000001-OH.csv', row.names = NULL)
oh_data <- subset(oh_data, select = -X)
# Categorize with party data
oh_data$party <- apply(oh_data['cand_nm'],1,rnParty)

oh_data_without_refund <- subset(oh_data, contb_receipt_amt > 0)


```

```{r Enrich data}
# Enrich with Month_Yr & Day of Month variables
oh_data$contb_receipt_dt <- as.Date(oh_data$contb_receipt_dt, "%d-%b-%y")
oh_data$Month_Yr <- format (as.Date(oh_data$contb_receipt_dt, "%d-%b-%y"), 
                            "%Y-%m")
oh_data$Month_Yr <- as.factor(oh_data$Month_Yr)
oh_data$Day_Month <- format (as.Date(oh_data$contb_receipt_dt, "%d-%b-%y"), 
                             "%d")
oh_data$Day_Month <- as.numeric(oh_data$Day_Month)

# Enrich with weekday variable
oh_data$contb_receipt_dt <- as.Date(oh_data$contb_receipt_dt,"%d-%b-%y")
oh_data$weekday <- weekdays(as.Date(oh_data$contb_receipt_dt))

oh_data$weekday <- factor(oh_data$weekday, levels= c("Monday", 
    "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

# Enrich with gender variable
# Step1. add a variable of contributor's surename
surname_split <- str_split_fixed(as.character(oh_data$contbr_nm) , ", ", 2)
colnames(surname_split) <- c("last name", "surname")
#surname <- surname_split[ , 2]
#typeof(surname_split)
#typeof(surname)
surname_split <-as.character(surname_split[ , "surname"])
surname <- str_split_fixed(surname_split, " ", 2)
oh_data$surname<- surname[, 1]

# Step2. 
gender_prediction <-gender(oh_data$surname, 
                           years = c(1897, 1997), # Voting age is 18+
                           method = "ssa")
gender_prediction <- gender_prediction[, c("name", "gender")]



oh_data$gender <- gender_prediction[
  match(oh_data$surname, gender_prediction$name), 
  "gender"]

```

## Univariate Plots Section

To know the basic about this dataset.  
```{r Univariate_Plot summary of the dataset and know the length of some key attributes}
cat("the total number of row in oh_data:", nrow(oh_data), "\n")
names(oh_data)
summary(oh_data)
```

#### By summary function to gain a general idea about the whole dataset
From the output and the definition of variables, I could know about the types of variables and decide the next exploration step.  

the total number of row in oh_data: 164475 rows
After enrichment, there are 24 variables.

The key questions I would like to anwser through this dataset are:  
1) if there is any correlation between contributed amount and the voting result?  
2) if there is any patterns for people donate funding? e.g. occupation, gender, city they live

**Key variable: ** donation amount (contb_receipt_amt, numeric variable)  
**other numeric variable for exploring distribution:** N/A  
**some important non-numeric variables:** candidate names(cand_nm), gender(gender), occupation(contbr_occupation), cities (contbr_city), party(party)
  

#### Distribution of donation amount
The distribution is quite spread and there are some negative numbers due to refund. For having a better view on donation amount, I used natural logarithm, log base 10, to transform my plot. With logarithm, I can see that the most common donation amount is around US\$50.1 (10^1.7) - US\$75(10^1.87)
  
```{r distribution of donation amount}
# Compute the number of classes for a histogram.
cat("Donated amount range: ", range(oh_data$contb_receipt_amt))
bin_number <- nclass.Sturges(oh_data$contb_receipt_amt)


u1 <- ggplot(oh_data, aes(x = contb_receipt_amt))+
  geom_histogram(bins = bin_number)

u1

# The distribution of contributed amount is wide and it seems to be skewed within the range of 0-500
ggplot(oh_data, aes(x = contb_receipt_amt)) +
  geom_histogram(binwidth = 10, boundary = 0) + # try to make a bin starting from 0
  geom_vline(aes(xintercept=mean(contb_receipt_amt, na.rm=T)), 
             color = "red", linetype = "dashed", size = 1) +
  coord_cartesian(xlim = c(-100, 500)) # zooming to the area of amount from -100 to $500

```
   
#### Logarithm transfermation of donation amount on the base of 10
I am using log base 10 for monetary amounts, because orders of ten seem natural for money: \$100, \$1000,\$10,000, and so on. The transformed data is easy to read.

  
```{r log_amount_distribution, cache=TRUE, cache.path = 'cache/', fig.path='figure/'}

u2 <- ggplot(oh_data, aes(x = contb_receipt_amt))+
  geom_histogram(bins = 100)+
  scale_x_log10()
u2

```

  
#### Basic idea on non-numeric variables: candidates, occupation, cities  
```{r Univariate_Plots: summary of the dataset and know the length of some key attributes}
# Length of key attributes
cat(length(unique(oh_data$cand_nm)), " unique candidates \n") # unique candidate
cat(length(unique(oh_data$contbr_occupation)), "unique occupations \n")
cat(length(unique(oh_data$contbr_city)), "unique contributed cities")

```

  
By plotting the bar charts and counting unique numbers of these non-numeric variables, there are too many unique data in terms of occupations and cities. It is difficult to read data from the graphs if ploting all occupations or cities so I plotted top 15 occupations and top 10 cities which contributed the most funding.

In terms of candidates, there are only 24 unique candidates so I used horizontal bar chart to show the full name of candidates. From the bar chart, Hiliary Clinton, Bernard Sanders and Donald Trump got the most contributed amount in Ohio state.  

Regarding to occupations, retired people contributed the most donation in Ohio. For city distribution, Columbas, Cincinnati and Clevenland are the top 3 cities to sponsor the election campaigns.
  
```{r Univariate_Plots: The distribution of every variable}
# candidate distribution

u3 <- ggplot(oh_data, aes( x = cand_nm )) +
  geom_bar(stat = "count") +
  coord_flip()

u3
# occupation distribution
# too many occupations to be seen on the graph

by_occupation <- group_by(oh_data, contbr_occupation)
by_occupation <- summarise(by_occupation, 
                           num_count = n())
by_occupation <- arrange(by_occupation, desc(num_count))
top_15_occupation <- head(by_occupation, 15)

u5 <- ggplot(top_15_occupation, aes(x = reorder(contbr_occupation, num_count), y = num_count)) +
  geom_bar(stat ="identity")+
  scale_x_discrete()+
  xlab("Occupations")+
  coord_flip()

u5

# city distribution
# too many cities to be seen on the graph
tab <- table(oh_data_without_refund$contbr_city)
tab_s <- sort(tab)
top10 <- tail(names(tab_s), 10)
d_s_top_city <- subset(oh_data, contbr_city %in% top10 )
d_s_top_city$contbr_city <- factor(d_s_top_city$contbr_city, levels = rev(top10))
head(d_s_top_city)

# Top 10 cities with the most number of contributions
u7 <- ggplot(d_s_top_city, aes(x = contbr_city)) +
  geom_bar()+
  theme_classic()+
  coord_flip()

u7
```


#### Month distribution
The election was on Nov. in 2016 but the donation started from March 2015 and reached the first peak on March 2016. The donation reached the highest peak on October 2016.


```{r Univariate_Plots: The distribution of contribution amount} 
# keep it for later

a <- ggplot(oh_data, aes( x = Month_Yr))

a + geom_histogram(stat = "count")+
  theme(axis.text.x = element_text(angle=60, hjust=1))



```
  
#### Party distribution
Although there are more donation records for Democratic party, there are more donated amount for Republican party. It might be caused by the average donation to Republican is higher. 
  


```{r Univariate_Plots: The number of contributions by party}
ggplot(oh_data, aes(x = party)) +
  geom_bar(stat = "count")


```
  
#### Gender distribution
The proportion of gender is almost equal (female : male is around 5 : 5)
  

```{r Univariate_Plots: gender distribution}

ggplot(oh_data, aes(x = gender)) + 
  geom_bar(stat = "count") +
  scale_x_discrete()


```

  
#### City level: donation records vs. donation amounts
I listed top 10 cities in terms of donation records and donation amounts. Take Columbus as an example, there are the most donation records among the cities but the donation amount is not the top 1 city. It shows that some cities might have more relatively small amount of donation.
  
```{r Univariate_Plots: Top 10 cities with the most amount of contributions,echo=FALSE}
# Top 10 cities with the most amount of contributions
by_amount <- group_by(oh_data, contbr_city)
city_by_amount <- dplyr::summarise(by_amount, 
          total_amount = sum(contb_receipt_amt)) 
city_by_amount <- arrange(city_by_amount, desc(total_amount))
city_by_amount
top_city <- head(city_by_amount, 10)
top_city
g2 <- ggplot(top_city, 
             aes(x = reorder(contbr_city, -total_amount), 
                 y= total_amount)) +
  geom_bar(stat = "identity") +
  coord_flip()

grid.arrange(u7, g2, nrow = 2)

```

  
  
# Univariate Analysis

### What is the structure of your dataset?
There are 164,475 obs in the Ohio dataset with 18 original varibles. For analysis purpose, I added 6 extra varibles (party, Month_Yr, weekday, day of month, surname and gender)  

### What is/are the main feature(s) of interest in your dataset?
The main features in the data set are "contb_receipt_amt" and the factors influencing the amounts. I’d like to find out which features have the most impact on raising more contributed amounts and I'd like to provide a few suggestions for candidates in the future when running a election found-raising campaign. I suspect city, occupation and day of week matter.  


### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
Since 2016 American presidential election result has came out, it would be great to do comparison analysis between contributed amount data and the final voting result data. I downloaded the voting result data for analyzing the correlation between contributed amount and the voters in Ohio. (The analysis is covered in the next section.)  

### Did you create any new variables from existing variables in the dataset?
Yes, I create 3 variables for further analysis. The 3 variables are listed below.  
1) Party: I categorized data into 3 categories(D, R, Other) based on candidate name  
2) Month_Yr: showing the contributed amount trend by month  
3) weekday: analyzing if there is a huge difference between weekday and weekend.  
4) Day_Month: the day of month
5) surname: for predicting the gender by gender library
6) gender: the gender of the contributors

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?
I enriched the Ohio dataset with Zipcode to visualize the contributed amount on Ohio map.(The analysis is conducted in multivariate plots section.)  

After merging with Ohio zipcode data from Zipcode library, I found there are 83 potential wrong zipcode data so I excluded them when I was plotting the contributed amount on the map. The reason why I excluded is that it is hard to identify the correct zipcode simply based on city names.  


# Bivariate Plots Section
#### First assumption: more donation amount means more votes  
Since the 2016 election result already came out, I enriched the original political finance dataset with vote data, which I found online[(link)]("https://raw.githubusercontent.com/tonmcg/County_Level_Election_Results_12-16/master/2016_US_County_Level_Presidential_Results.csv"). Since I only analyzed Ohio data, I picked out the Ohio voter data from the national vote data by using subset and mapped with my original contribution amount data by city name. (which in the vote data, the column called "county_name")

Regarding to the mapping, there are different format of city names. There is "county" at the end of each city name in vote data so I have to trim the word of "county". The cases of letter in these 2 data frames are different as well. To properly map the data, I chang the cases of letter by the function of "tolower".



```{r Bivariate_Plots: Coorelation matrix plot, echo=FALSE, message=FALSE, warning=FALSE}

# names of columns to download data
vote_col <- c("county_name", 
              "votes_dem", 
              "votes_gop", 
              "total_votes", 
              "state_abbr") 

# Load vote data from the online source
vote_result <- fread("https://raw.githubusercontent.com/tonmcg/County_Level_Election_Results_12-16/master/2016_US_County_Level_Presidential_Results.csv", select = vote_col)

simple_oh_vote <- subset(vote_result,  state_abbr == "OH")

# Clean Ohio vote data
split_function <- str_split_fixed(simple_oh_vote$county_name, " ", 2) # take off county
simple_oh_vote$county_name <- sapply(split_function, "[", 1)
simple_oh_vote$county_name <- tolower(simple_oh_vote$county_name)

#Merge vote data with contributed data
by_city <- group_by(oh_data, contbr_city, party)
by_city <- dplyr::summarise(by_city, 
                     count = n(),
                     total_amount = sum(contb_receipt_amt))
by_city$contbr_city <- tolower(by_city$contbr_city)
print(by_city)
by_city.wide <- dcast(by_city, 
                      contbr_city ~ party, 
                      value.var = 'total_amount')
by_city.wide[is.na(by_city.wide)] <- 0
head(by_city.wide)


New_by_city <- merge(by_city.wide, simple_oh_vote, 
                     by.x = 'contbr_city', by.y = 'county_name')

colnames(New_by_city) <- c("contbr_city", 
                           "amount_D", 
                           "amount_Other", 
                           "amount_R", 
                           "votes_D",
                           "votes_R",
                           "total_votes",
                           "state_abbr")
New_by_city$total_amount <- New_by_city$amount_D + 
  New_by_city$amount_R+ New_by_city$amount_Other

New_by_city$state_abbr <- NULL

print ("New dataset of donation amount and votes")
head(New_by_city)

```
  
#### Checking the correlation between donation amount and votes

```{r Corrplot_amount_vote, cache=TRUE, cache.path = 'cache/', fig.path='figure/'}
N <- New_by_city
rownames(N) <- N$contbr_city
N$contbr_city <- NULL
N <- cor(N)
cor_plot <- corrplot(N, method="number", is.corr=FALSE)

```
  
#### Finding: the correlation between donation amount and votes is not strong as I expected.
A minor finding might be that the correlation between donation amount and votes is stronger for Republican supporters (the correlation corefficient: 0.4)

```{r}

r1 <- ggplot(data = New_by_city, aes(x= total_amount, y= total_votes), 
             group = contbr_city)+
  geom_point()+
  geom_smooth(method = "lm", se= FALSE)+
  ggtitle("Relationship: Total Contributed Amount vs. Total Votes")+ 
  scale_y_continuous(labels = comma)
r1 


```

So I tried to exclude outliers by limiting x-axis and y-axis to focus on the bulk data below. The slope for Plot1-2 (Republican party data) is slightly steeper than Plot1-3(Democratic party).  
```{r Excluding_outliers}

r1 <- ggplot(data = New_by_city, aes(x= total_amount, y= total_votes), 
             group = contbr_city)+
  geom_point()+
  geom_smooth(method = "lm", se= FALSE)+
  ggtitle("Plot1-1. Relationship between Total Contributed Amount and Total Votes")+ 
  scale_y_continuous(labels = comma, limits = c(0, 200000))+
  scale_x_continuous(limits = c(0, 30000))


r2 <- ggplot(data = New_by_city, aes(x= amount_R, y= votes_R), 
             group = contbr_city)+
  geom_point()+
  geom_smooth(method = "lm", se= FALSE)+
  ggtitle("Plot1-2. Relationship: Contributed Amount vs. Votes of Republican Party")+ 
  scale_y_continuous(labels = comma, limits = c(0, 200000))+
  scale_x_continuous(limits = c(0, 30000))


r3 <- ggplot(data = New_by_city, aes(x= amount_D, y= votes_D), 
             group = contbr_city)+
  geom_point()+
  geom_smooth(method = "lm", se= FALSE)+
  ggtitle("Plot1-3. Relationship: Contributed Amount vs. Votes of Democratic party")+ 
  scale_y_continuous(labels = comma, limits = c(0, 200000))+
  scale_x_continuous(limits = c(0, 10000))

grid.arrange(r1, r2, r3, nrow = 3)
```


#### Exploring the donation amount trend by month
Based on the line chart below, the wave of donation starts from late July in 2015. There might be some interesting insights to be explored.

```{r Bivariate_Plots_amount_change_by_month, echo=FALSE, cache=TRUE, cache.path = 'cache/', fig.path='figure/'}

by_month <- group_by(oh_data, Month_Yr)
by_month <- summarise(by_month, 
                      count = n(),
                      amount = sum(contb_receipt_amt))

ggplot(oh_data, aes(x = Day_Month, y = contb_receipt_amt, group = 1)) +
  geom_line(stat = "summary", fun.y = sum)+
  scale_x_continuous(breaks = seq(1,31,10))+
  scale_y_continuous(labels = dollar_format("US$"))+
  facet_wrap(~Month_Yr, ncol= 5)+
  theme(axis.text.x = element_text(size=8)) 



```

#### Donation trend by weekday
The intuition is that there might have more donation during weekend.  

But surprisingly, there are more donation on weekday. I guess it might be the different lifestyle between Asia and United States.

```{r Bivariate_Plots: the day of week v.s. Contributed amount,echo=FALSE}
oh_data$contb_receipt_dt <- as.Date(oh_data$contb_receipt_dt,"%d-%b-%y")
oh_data$weekday <- weekdays(as.Date(oh_data$contb_receipt_dt))

oh_data$weekday <- factor(oh_data$weekday, levels= c("Monday", 
    "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

by_weekday <- group_by(oh_data, weekday)
by_weekday_sum <- summarise(by_weekday, 
                        count = n(),
                        amount = sum(contb_receipt_amt))
by_weekday <- group_by(oh_data, weekday)
by_weekday_mean <- summarise(by_weekday, 
                        count = n(),
                        amount = mean(contb_receipt_amt))

w1 <- ggplot(by_weekday_sum, aes( x = weekday, y = amount, group =1)) +
  labs(y = "Total amount")+
  geom_bar(stat = "identity")+
  ggtitle("Sum of contributed amount by weekday")+ 
  scale_y_continuous(labels = dollar_format("US$"))

w2 <- ggplot(by_weekday_mean, aes( x = weekday, y = amount, group =1)) +
  labs(y = "Average amount")+
  geom_bar(stat = "identity")+
  ggtitle("Average of contributed amount by weekday")+ 
  scale_y_continuous(labels = dollar_format("US$"))


grid.arrange(w1, w2, nrow =2)
```


# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?
I noticed that the relationship between distributed amount and the number of voters is not positively strong. It seems to have week relationship which is against my original assumption.

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?
When dicussing the relationship between the contributed amount and the toal voters, Republican party supporters show stronger correlation than Democratic party supporters.

The correlation coefficient between contributed amount and voter numbers
1)Republican party : 0.401
2)Democratic party : 0.184

The correlation coefficient is higher than the correlation coefficient of total contributed amount and total voter numbers in Ohio (which is 0.307)

### What was the strongest relationship you found?
The relationship between the total contributed amount and the contributed amount of Republican party is super relative (the correlation coefficient is 0.934) because the contributed amount from Republican party supporters accounts for ~60%. 

However, this is not a proper pair to check the relationship because these 2 factors are not independent. 


# Multivariate Plots Section

  
#### Donation by occupation
In order to increase more donation amount, I would also like to find out which occupation contributed the most in terms of the number of donation and the average amount which they donated each time. 

Based on the graph below, I could see that "Attorney" and "Homemaker" have the higher average donation amount. It provides the direction of party platform which might need to be beneficial for these 2 occupations.  

```{r Multivariate_Plots: analysis by occupation - we may try to increase the average amount, echo=FALSE}
by_occupation <- group_by(oh_data, contbr_occupation)
by_occupation_amount <- summarise(by_occupation,
                               count = n(),
                               average_amount = mean(contb_receipt_amt),
                               total = sum(contb_receipt_amt)
)

ordered_data <- arrange(by_occupation_amount, desc(count))
top10_occupation <- head(ordered_data, 10)

print("Top10 occupation:")
print(top10_occupation$contbr_occupation)


ggplot(top10_occupation, 
       aes(x = reorder(contbr_occupation, -count), y = count))+
  geom_bar(stat = "identity")+
  geom_line(data = top10_occupation, aes(y = average_amount), 
            colour = "blue", group =1 )+
  scale_y_continuous(sec.axis = sec_axis(~./400, name = "Average Amount"))


# two plots
p1 <- ggplot(top10_occupation, 
             aes(x = reorder(contbr_occupation, -count), y = count)) +
  geom_bar(stat = "identity")  + theme_bw()+
  scale_x_discrete(labels = abbreviate)+ 
  scale_y_continuous(labels = comma)
p2 <- ggplot(top10_occupation, 
             aes(x = reorder(contbr_occupation, -count), y = count)) +
geom_line(data = top10_occupation, aes(y = average_amount), 
          colour = "blue", group =1 ) +
  scale_x_discrete(labels = abbreviate) + 
  scale_y_continuous(labels = dollar_format("US$"), name = "Average donation amount") +
  theme_bw() %+replace% 
  theme(panel.background = element_rect(fill = NA))


# extract gtable
g1 <- ggplot_gtable(ggplot_build(p1))
g2 <- ggplot_gtable(ggplot_build(p2))

# overlap the panel of 2nd plot on that of 1st plot
pp <- c(subset(g1$layout, name == "panel", se = t:r))
g <- gtable_add_grob(g1, g2$grobs[[which(g2$layout$name == "panel")]], 
                     pp$t, pp$l, pp$b, pp$l)

# axis tweaks
ia <- which(g2$layout$name == "axis-l")
ga <- g2$grobs[[ia]]
ax <- ga$children[[2]]
ax$widths <- rev(ax$widths)
ax$grobs <- rev(ax$grobs)
ax$grobs[[1]]$x <- ax$grobs[[1]]$x - unit(1, "npc") + unit(0.15, "cm")
g <- gtable_add_cols(g, g2$widths[g2$layout[ia, ]$l], length(g$widths) - 1)
g <- gtable_add_grob(g, ax, pp$t, length(g$widths) - 1, pp$b)

# draw it
grid.draw(g)
```
  
#### Checking the contribution by cities

During the election period, candidates need to travel around the cities to win supports. Using the heat map would provide a clear visualization on where the most supportive cities in terms of donation amount. I enriched the original dataset with zipcode data to plot a heatmap afterwards.  

```{r Merging zipcode data with original dataset, echo=FALSE}

# subset data for zipcode
cNamelist <- c("cand_nm", 
               "contbr_city", 
               "contbr_zip", 
               "contb_receipt_amt", 
               "party")
idx <- which(colnames(oh_data) %in% cNamelist)
NewDF <- oh_data[, idx]
head(NewDF)

# Merge zipcode data with original dataset
data(zipcode)
NewDF$contbr_zip <- clean.zipcodes(NewDF$contbr_zip)
NewDF <- merge(NewDF, zipcode, by.x = 'contbr_zip', by.y = 'zip')

```

#### Check potential error data after merging zipcode data into original dataset below.
```{r donation_heatmap, cache=TRUE, cache.path = 'cache/', fig.path='figure/'}
# Check potential error data: found 83 potential wrong zipcode data
potential_mistake <- subset(NewDF, state != 'OH')
cat("Potential error data: ", nrow(potential_mistake))



# Exclude data with wrong zipcode(outside of Ohio state)
map<-get_map(location='ohio state', zoom=7, maptype='roadmap')
ggmap(map)+
  geom_point(aes(x=longitude, y=latitude, size=(contb_receipt_amt), color = party), 
             data=subset(NewDF, state == "OH"), alpha=1/50) +
  scale_color_manual(name = "Party",
  labels = c("Democratic", "Other", "Republican"),
  values = c("red", "grey", "blue")) +
  theme(legend.position = "bottom") +
  guides(size=FALSE, colour = guide_legend(override.aes = list(alpha = 1)))


```

#### Checking the contribution by cities

Checking the relation between time and donation amount in top 10 cities, I could see an intersting fact that the donation amount supporting Democratic party skewwed in 2016. On the other hand, the donation to Republican party skewwed in 2015. Cincinnati and Cleveland, the two cities with the most donation amount, have the most obviouse trend.  

```{r Multivariate_Plots: donation_by_month_by_party, echo=FALSE, cache=TRUE, cache.path = 'cache/', fig.path='figure/'}

# 1. Plot charts
top1 <- ggplot(subset(d_s_top_city, party == "Democratic"), 
               aes(x=Month_Yr, y=contb_receipt_amt, group= contbr_city))+
  geom_line(aes(color = contbr_city), stat = "summary", fun.y=sum)+
  scale_y_continuous(labels = dollar_format("US$"))+
  theme(axis.text.x = element_text(angle=60, hjust=1))+
  ggtitle("Month-Year vs. Donation Amount(Democratic)")+
  theme(legend.position="bottom")
  

top2 <- ggplot(subset(d_s_top_city, party == "Republican"), 
               aes(x=Month_Yr, y=contb_receipt_amt, group= contbr_city))+
  geom_line(aes(color = contbr_city), stat = "summary", fun.y=sum)+
  scale_y_continuous(labels = dollar_format("US$"))+
  theme(axis.text.x = element_text(angle=60, hjust=1))+
  ggtitle("Month-Year vs. Donation Amount(Republican)")+
  theme(legend.position="none")

# 2. Save the legend
legend <- get_legend(top1)
# 3. Remove the legend from the box plot
top1 <- top1 + theme(legend.position="none")

grid.arrange(top1, top2, legend, nrow = 3, heights = c(8,8,2))



```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?
I noticed that the major cities account for more contributed amount. After visualing on the map, it shows clearly that there are a few of heat spots in Ohio.  

Checking the relation between time and donation amount in top 10 cities, I could see an intersting fact that the donation amount supporting Democratic party skewwed in 2016. On the other hand, the donation to Republican party skewwed in 2015. Cincinnati and Cleveland, the two cities with the most donation amount, have the most obviouse trend.


### Were there any interesting or surprising interactions between features?
After distinguishing the contributed amount by party, it shows that there are more funding going to Republican party and it refelects on voting result that Republican party won Ohio at the end.


### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.
No. I tried to build a linear regression model between numeric and catergorical data but it failed and it seems to involve more complexing statistical library.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

grid.arrange(r1, r2, r3, nrow = 3)
```

### Description One
The correlation between donation amount and votes is not as strong as I expected.
From the previous exploration, I found there are a lote of cities skewwed together, So I tried to exclude outliers by limiting x-axis and y-axis to focus on the bulk data. The slope for Plot1-2 (Republican party data) is slightly steeper than Plot1-3(Democratic party).

A minor finding might be that the correlation between donation amount and votes is stronger for Republican supporters (the correlation corefficient: 0.4)  
  
The correlation coefficient between contributed amount and voter numbers for 2 parties.  
1)Republican party : 0.401 (plot 1-2)  
2)Democratic party : 0.184 (plot 1-3)  

The correlation coefficient is higher than the correlation coefficient of total contributed amount and total voter numbers in Ohio (which is 0.307, plot 1-1)  


### Plot Two
```{r echo=FALSE, Plot_Two}
grid.arrange(top1, top2, legend, nrow = 3, heights = c(8,8,2))

```

### Description Two
Based on the previous exploration of contributed amount by weekday, it shows that there is lower contributed amount on weekend. The weekday finding is somewhat telling about weekday vs weekend habits but this might be expected. So I diveed into looking at contribution amounts on a broader time scale.  

Checking the relation between time and donation amount in top 10 cities, I could see an intersting fact that the donation amount supporting Democratic party skewwed in 2016. On the other hand, the donation to Republican party skewwed in 2015. Cincinnati and Cleveland, the two cities with the most donation amount, have the most obviouse trend.  

My guess is that the different distrinution might be caused by the party platform annoucements of each party or the campaign tour plan. I think the donation amount should be boosted everytime when a party's candidate engages with citizen in a city.   

### Plot Three
```{r Plot_Three, echo=FALSE, cache=TRUE, cache.path = 'cache/', fig.path='figure/'}
map_total <- ggmap(map)+
  geom_point(aes(x=longitude, y=latitude, size=contb_receipt_amt), 
             data=subset(NewDF, state == "OH"), alpha=1/150, 
             color = "lightgreen")+
  theme(legend.position="none")+
  ggtitle("Plot3. Contributed amount density")+
  guides(size=FALSE)

map_party <- ggmap(map)+
  geom_point(aes(x=longitude, y=latitude, size=contb_receipt_amt, 
                 color = party), 
             data=subset(NewDF, state == "OH"), alpha=1/50)+ 
  scale_colour_manual(values=c("Republican" = "pink", "Democratic" = "blue", "Other" = "grey"))+
  theme(legend.position = "right")+
  ggtitle("Plot3-1. Amount density by party")+
  guides(size=FALSE, colour = guide_legend(override.aes = list(alpha = 1)))

grid.arrange(map_total, map_party, ncol= 2)


```

### Description Three
It shows that the contributed money is mainly from city area such as Columbus, Cleveland, Akron and Cincinnati etc. It helps candidates to identify the cities to plan their future campaigns for raising more funding.


I distinguish the funding for Republican party and Democratic party by color in Plot3-1. It shows that there are more funding for Republican party in Ohio and the voting result also shows that Republican party won Ohio state. 


------

# Reflection
Before starting the analysis, I assumed that the contributed amount would be a strong indicator for election result. After analyzing the relationship between the election result of Ohio and the contributed amount data of Ohio. The correlation coefficient between these 2 factors are lower than I expected and it can't be suspected as having strong correlation between contributed amount and voter numbers.

However, this is only analyzing one state. I think, for optimizing/ further analayzing, I would suggest to analyze the data of all states in the U.S. to see if there are any strong relationship between these 2 factors. 

During the analysis, I was quite struggling with more than 6,000 occupations which I thought there might be some insights to br cracked. It would be better if there are some default options for people to choose while they are making donation, such as "Retired", "Public Servant", "Military Soldiers" or "Teachers" etc. I could cross-check with each party's party platform to see if party platform have any impact on donation amounts by occupation.   

### Problem encountered and the solution records
**Problem: Error: Discrete value supplied to continuous scale all the time.**  
After merging the predicting gender by gender library, I got error message while I tried to plot a bar chart of gender with ggplot2. While I used ggplot(oh_data, aes(x = gender)), I got "Error: Discrete value supplied to continuous scale all the time".  

One way to correct Discrete value supplied to continuous scale all the time is by plotting discrete variable on a discrete scale. There are a few ways to do this:  
1) Add scale_x_discrete() layer. --> this works!  
2) use as.factor() i.e. ggplot(oh_data, aes(x=as.factor(gender))) + geom_bar() --> this didn't work on my code. I got another error: *Error in sort.list(y) : 'x' must be atomic for 'sort.list' Have you called 'sort' on a list?*  

  
### Reference articles or links to solve problems
[Log Transformations for Skewed and Wide Distributions](https://www.r-statistics.com/2013/05/log-transformations-for-skewed-and-wide-distributions-from-practical-data-science-with-r/)

[An Introduction to corrplot Package](https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html)

[ggplot2 axis ticks : A guide to customize tick marks and labels](http://www.sthda.com/english/wiki/ggplot2-axis-ticks-a-guide-to-customize-tick-marks-and-labels)

[Compute the number of classes for a histogram.](https://stat.ethz.ch/R-manual/R-devel/library/grDevices/html/nclass.html)

[Problem while loading data: duplicate 'row.names' are not allowed error in R programming](https://stackoverflow.com/questions/8854046/duplicate-row-names-are-not-allowed-error-in-r-programming/18966905#18966905)

[Why does a boxplot in ggplot requires axis x and y?](http://stackoverflow.com/questions/25049104/why-does-a-boxplot-in-ggplot-requires-axis-x-and-y)
[What does stat means in ggplot?](https://discuss.analyticsvidhya.com/t/what-does-stat-means-in-ggplot/6579)
[ggplot2 line chart gives “geom_path: Each group consist of only one observation. Do you need to adjust the group aesthetic?”](http://stackoverflow.com/questions/27082601/ggplot2-line-chart-gives-geom-path-each-group-consist-of-only-one-observation)
[Find the day of a week in R](http://stackoverflow.com/questions/9216138/find-the-day-of-a-week-in-r)

[Creating a Density Map in R with Zipcodes](http://www.austinwehrwein.com/digital-humanities/creating-a-density-map-in-r-with-zipcodes/)

[2016_US_County_Level_Presidential_Results.csv](https://github.com/tonmcg/County_Level_Election_Results_12-16/blob/master/2016_US_County_Level_Presidential_Results.csv)